---
title: "Competitiveness Analysis"
output:
  html_document:
    df_print: paged
---
The broad goal is to measure if the introduction of an independent redistricting commission in California had an impact on competitiveness of the races. 

#Get data

This is an [State Leglislative electoral data](https://github.com/PrincetonUniversity/historic_state_legislative_election_results/tree/2bf28f2ac1a74636b09dfb700eef08a4324d2650) gathered by Will Adler. When you execute code within the notebook, the results appear beneath the code. 

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
# state legislative 
data = read_csv("https://raw.githubusercontent.com/PrincetonUniversity/historic_state_legislative_election_results/2bf28f2ac1a74636b09dfb700eef08a4324d2650/state_legislative_election_results_post1971.csv")
# filter to ca post-2000
stateleg = data %>%
  filter(State == "CA", Year > 1999)

# use this to turn on/off imputation
# can also try filtering out uncontested races
impute = TRUE

stateleg %<>% mutate(`D Voteshare` = ifelse(`D Voteshare` == 1, 0.75, `D Voteshare`),
                     `D Voteshare` = ifelse(`D Voteshare` == 0, 0.25, `D Voteshare`))

stateleg %<>% mutate(pre_2011 = ifelse(Year < 2012, "Pre-2011", "Post-2011"),
                     R_voteshare = 1 - `D Voteshare`,
                     win_margin = ifelse(R_voteshare > 0.5, R_voteshare - 0.5, `D Voteshare` - 0.5))
head(stateleg)


# state senate - return to do this

# state/decade specific data
D_pre2011 = stateleg %>%
  filter(Party == "D") %>%
  filter(pre_2011 == "Pre-2011")

D_post2011 = stateleg %>%
  filter(Party == "D") %>%
  filter(pre_2011 == "Post-2011")

R_pre2011 = stateleg %>%
  filter(Party == "R") %>%
  filter(pre_2011 == "Pre-2011")

R_post2011 = stateleg %>%
  filter(Party == "R") %>%
  filter(pre_2011 == "Post-2011")

```

#Pool Data
 
We pool together 10 years of data in order to calculate the standard deviation of the voteshare: pre-2011 and post-2011. We do this in order to investigate the variability of the voteshare over time. We also check the average win margin and standard deviation of the win margins broken down by decade and party. 

```{r}
table(stateleg$pre_2011)
stateleg %>% 
  group_by(pre_2011) %>%
  summarise(n_obs = n(), 
            mean_vs = mean(`D Voteshare`), 
            sd_vs = sd(`D Voteshare`),
            mean_win_margin = mean(win_margin),
            sd_win_margin = sd(win_margin))

stateleg %>% 
  group_by(pre_2011, Party) %>%
  summarise(mean_winmargin = mean(win_margin),
            sd_winmargin = sd(win_margin))

hist(stateleg$win_margin)
# D wins win margin
# pre-2011
hist(D_pre2011$win_margin)

# post-2011
# way more uncontested races for D's post-2011
hist(D_post2011$win_margin)

# R wins win margin 
# pre-2011
hist(R_pre2011$win_margin)

# post-2011
hist(R_post2011$win_margin)

```
**(Without imputing uncontested races)**
Interesting! In both decades the average win margin was larger for Democrats than Republicans, and the marginal increase was about equal for both from pre-2011 to post-2011 (3.6 increase for R, 4.2% increase for D). This suggests that competitiveness decreased for both from pre-2011 to post-2011, and got marginally worse for Democrats. However, the variability increased for both parties from pre-2011 to post. 

So without looking much further, it seems like post-2011 things got more competitive for Republicans and less competitive for Democrats. :0

**(With imputing uncontested races)**
Things look different when we impute unconsted races (100%/0%) to (75%/25%).The average win margins don't change much within party across the time period. For Republicans, the win margin is consistent at ~13% and for Democrats, the win margin is consistent at ~19-20%. 

The results are more variable for Democrats than Republicans in both periods. This suggests that the redistricting didn't make a huge difference in competitiveness for either party. 

```{r include=FALSE}
# don't think this actually tells us too much because the districts are different across decades
small_multiples = ggplot(data=stateleg, aes(x=Year, y =`D Voteshare`, colour = pre_2011, group = pre_2011)) + 
    geom_rect(aes(xmin = -Inf, 
                  xmax = Inf, 
                  ymin = 0.45, 
                  ymax = 0.55), 
            alpha = 0.5,fill = "grey", colour="white") + 
      geom_line() +
  facet_wrap(~District)
small_multiples
#ggsave('small_multiples.pdf', small_multiples, width=8, height=8)
```

Next, want to calcualte the win margin in a given year. We do this for both by party and irregardless of party.

```{r}
stateleg$winning_vs = ifelse(stateleg$`D Voteshare` > 0.5, stateleg$`D Voteshare`, stateleg$R_voteshare)

stats_out <- stateleg %>%
  group_by(pre_2011) %>%
  summarise(mean_wins = mean(winning_vs),
            sd_wins = sd(winning_vs)) %>%
  mutate(wins_1sigma = mean_wins + sd_wins,
         wins_2sigma = mean_wins + 2*sd_wins)
stats_pre = stats_out %>%
  filter(pre_2011 == "Pre-2011")
stats_post = stats_out %>%
  filter(pre_2011 == "Post-2011")

stateleg %<>% mutate(competitive_1sigma = ifelse(ifelse(pre_2011 == "Pre-2011", 
                                                        (winning_vs > stats_pre$wins_1sigma), 
                                                        (winning_vs > stats_post$wins_1sigma)), 
                                                 "No", "Yes"),
                     competitive_2sigma = ifelse(ifelse(pre_2011 == "Pre-2011", 
                                                        (winning_vs > stats_pre$wins_2sigma), 
                                                        (winning_vs > stats_post$wins_2sigma)), 
                                                 "No", "Yes"))

table(stateleg$competitive_1sigma)
table(stateleg$competitive_2sigma)

comp_2sigma = stateleg %>% filter(competitive_2sigma == "Yes")
nocomp_2sigma = stateleg %>% filter(competitive_2sigma == "No")

hist(comp_2sigma$win_margin)
hist(nocomp_2sigma$win_margin)

```

Try it with imputing uncontested races to 75%/25%!


Aggregate population totals from districts - opportunity to elect (not district-specific, but on aggregate)















